Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-06-02 ms.tex
Modify Section 5
From f6de6ce:
============

In ms.tex,

(*) Modify Section 5 to emphasize the need for detailed study of
coupling between beam response near horizon and foreground
contamination

The delay spectrum for a baseline vector, $\boldsymbol{b}$, is \citep[][Paper~I]{par12a,par12b,thy13}: 
\begin{align}
  \widetilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
with interferometer visibilities, $V_b(f)$, given by \citep{van34,zer38,tho01}:
\begin{align} 
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega \\
         &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, as a function of frequency ($f$) and direction on the sky denoted by the unit vector $\hat{\boldsymbol{s}}\equiv (l,m,n)$, $W_\textrm{i}(f)$ denotes instrumental bandpass weights, $W(f)$ is a spectral weighting function that controls the transfer function in the delay transform, $\dif\Omega=(1-l^2-m^2)^{-1/2}\,\dif l\,\dif m$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector, and $c$ is the speed of light. $\tau=\boldsymbol{b}\cdot\hat{\boldsymbol{s}}/c$ is the geometric delay between antenna pairs measured relative to the zenith and provides a mapping to position on the sky.

@@

Delay spectrum maps the geometric delays to positions of foreground objects on the sky. Thus the directional power pattern of the antenna has a direct impact on the delay spectrum. Since the contamination in the {\it EoR window} is strongly dependent on sources of emission close to the horizon, design of antenna apertures that suppress their sensitivity towards the horizon is important.

We approached this issue in Paper~I via simulations of different antenna apertures, namely, a dipole, a phased array, and a dish. The relative foreground contamination in the {\it EoR window} was found to be roughly $10^4:10^2:1$ respectively. In addition, the foreground contamination inside the {\it foreground wedge} was found to be suppressed by a factor $10^4$ for a dish relative to an MWA-like phased array thus raising the possibility of accessing the EoR signal in $k$-modes inside the {\it foreground wedge}.

The response of new instruments such as SKA and HERA to foreground contamination has to be investigated thoroughly. With the SKA at low frequencies, plans to deploy ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas are being considered. The ``station beam'' will be obtained by a phased addition of the dipole responses. The typical horizon response of these beams is $\gtrsim$~30~dB lower than at zenith\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.} but the randomness across different stations in the placement of dipoles is expected to reduce the response significantly further as pairs of station voltage beams are multiplied with each other. HERA will be a closely packed hexagonal array of fixed 14~m dishes which will observe the sky drifting overhead with redundant antenna spacings. It will use dishes that limit response near the horizon to $\lesssim -40$~dB, $\sim$~20~dB lower than that of the MWA. Their precise beam responses and the resulting foreground contamination, especially near the horizon, will prove to be critical and timely inputs to their aperture designs which are in their final stages.

>>>>

The delay spectrum for a baseline vector, $\boldsymbol{b}$, is \citep[][Paper~I]{par12a,par12b,thy13}: 
\begin{align}\label{eqn:delay-spectrum}
  \widetilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
with interferometer visibilities, $V_b(f)$, given by \citep{van34,zer38,tho01}:
\begin{align} 
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega \label{eqn:vis1}\\
         &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, \label{eqn:vis2}
\end{align}
where, $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, as a function of frequency ($f$) and direction on the sky denoted by the unit vector $\hat{\boldsymbol{s}}\equiv (l,m,n)$, $W_\textrm{i}(f)$ denotes instrumental bandpass weights, $W(f)$ is a spectral weighting function that controls the transfer function in the delay transform, $\dif\Omega=(1-l^2-m^2)^{-1/2}\,\dif l\,\dif m$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector, and $c$ is the speed of light. $\tau=\boldsymbol{b}\cdot\hat{\boldsymbol{s}}/c$ is the geometric delay between antenna pairs measured relative to the zenith and provides a mapping to position on the sky.

@@

Delay spectrum maps the geometric delays to positions of foreground objects on the sky. Thus the directional power pattern of the antenna has a direct impact on the delay spectrum. In fact, from equations~\ref{eqn:delay-spectrum}, \ref{eqn:vis1} and \ref{eqn:delay-power-spectrum}, the delay power spectrum scales as the square of the directional power pattern of the antenna. Since the contamination in the {\it EoR window} is strongly dependent on sources of emission close to the horizon, design of antenna apertures that suppress their sensitivity towards the horizon is important.

We approached this issue in Paper~I via simulations of different antenna apertures, namely, a dipole, a phased array, and a dish. Among these, a dish provides the best quality as evaluated from a foreground contamination viewpoint. With new instruments such as HERA and SKA on the horizon, we extend this discussion to further emphasize the need to comprehensively study the effect of their antenna apertures on foreground contamination.

% The relative foreground contamination in the {\it EoR window} was found to be roughly $10^4:10^2:1$ respectively. In addition, the foreground contamination inside the {\it foreground wedge} was found to be suppressed by a factor $10^4$ for a dish relative to an MWA-like phased array thus raising the possibility of accessing the EoR signal in $k$-modes inside the {\it foreground wedge}.

% The response of new instruments such as SKA and HERA to foreground contamination has to be investigated thoroughly.

HERA will deploy 14~m dishes with 331 of them in a closely packed hexagonal array and another 21 as outliers. With the advantage of enormous redundancy and unprecedented sensitivity, it will address the following key questions -- what objects first lit up the Universe and reionized the neutral IGM? Over what redshift range did this occur? And how did the process proceed, leading to the large scale galaxy structure seen today? The fixed dishes will observe the sky drifting overhead. Based on simulations of uniformly illuminated circular disks, the response near the horizon is $\lesssim -40$~dB, $\gtrsim$~20~dB lower than that of the MWA. Tapering the reflectivity of the dish is being studied to further reduce this response. The responses of the actual dishes will be tested shortly.

With the SKA at low frequencies, deploying ``stations'' of $\sim$~35~m diameter that act as aperture arrays consisting of 256 pseudo-randomly placed vertical log-periodic dipole antennas is being considered\footnote{Documents at \url{https://www.skatelescope.org/key-documents/} on SKA re-baselining}$^,$\footnote{SKA memo ``Station Response and Imaging Performance of LFAA: 100 MHz to 600 MHz'' by Razavi-Ghods et al.}. The voltage beam of the station will be obtained by a phased addition of the dipole responses. Since the power pattern of a baseline consisting of two stations will be effectively a product of two voltage beams, it will tend to homogenize the fluctuations in the power patterns of individual stations. Even with unconstrained, and hence sub-optimal, realizations of these phased array stations, we estimate the typical horizon response of these beams to be $\gtrsim$~30--40~dB lower than at zenith, also $\sim 20$~dB lower than the MWA beam. Constrained optimization studies of antenna layouts in the stations are under active study and could yield even better responses.

From these simple antenna aperture models, HERA and SKA should mitigate foreground contamination by $\gtrsim 40$~dB relative to the MWA. The precise beam responses and the resulting foreground contamination, especially near the horizon, will prove to be critical and timely inputs to the actual aperture designs of HERA dishes and SKA stations which are nearing their final stages.
